\section{Related Work}
\label{sec:related-work}

To the best of our knowledge, we are the first to provide a generic
operational framework for the characterization of CRDT specifications
based on the axiomatic characterizations of~\cite{Burckhardt:2014}
under diverse delivery assumptions of the underlying network.
%

There have been many
efforts~\cite{Burckhardt:2014,Burckhardt:2014b,ZellerBP14,MukundRS15}
in formalizing the semantics of replicated data types.
%
Unlike the approach in~\cite{Burckhardt:2014,Burckhardt:2014b}, our
proofs are based on standard simulation arguments, common to proofs in
programming languages.
%
Moreover, unlike~\cite{Burckhardt:2014,Burckhardt:2014b}, we provide
operational variants of the specifications which are instrumental in
proving the correctness of implementations.

In~\cite{ZellerBP14}, the proofs provided are based on state-based
implementations of CRDTs, our work is concerned with operation-based
implementations.

An important difference between our work and the aforementioned is
that to be able to express a simulation relation, we need to export at
the level of the operational specification the delivery assumptions of
the system.
%
In turn this means that our proofs are with respect to a more concrete
specification than these other works, since the delivery assumptions
are not apparent in their specifications.
%
We will consider in future work how to prove that out reference
semantics preserves the specifications \'a la~\cite{Burckhardt:2014}
where the delivery conditions have been removed.
%

Perhaps the work that is closer to ours is~\cite{MukundRS15,MukundRS15b} 
since they also use LTS-based specifications. 
Unlike those works, our specifications of CRDTs are given in the
format of~\cite{Burckhardt:2014} and the reference implementing is
obtained thereof.
%
Moreover, while they propose a CEGAR methodology to carry out the
proofs of the data type implementations, our proofs are based on
simulation arguments.
%
Finally, they do not provide proofs of concrete algorithms in their
work. 

There are works on the verification of program properties about
Credits~\cite{GotsmanYFNS16}.
%
We consider our work to be orthogonal~\cite{GotsmanYFNS16} since we
are proving the correctness of implementations w.r.t. the
specification. 

\section{Conclusion}
\label{sec:conclusion}

We have developed a methodology to prove CRDT implementations correct
w.r.t. their specifications given in the axiomatic style
of~\cite{Burckhardt:2014}.
%
Our methodology constructs a reference implementation of the
specification where we make apparent the assumptions of message
delivery of the underlying network. 
% 
Based on these references implementations of the specification we
provide simulation arguments to show that concrete implementations of
the CRDT meet the specification, all the while taking into
consideration the message delivery assumed by the specification and
the implementation.
%
We show the practicality of our methodology by proving the correctness
of RGA~\cite{RohJKL11} and two different OR-set
implementations~\cite{Shapiro:2011,Bieniusa:2012}.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft"
%%% End:
