%!TEX root = draft.tex

\section{Simulation Relation for CRDT Implementations}
\label{sec:simulation relation for CRDT implementations}


\subsection{Collection Specification with Causal Delivery}
\label{subsec:collection specification with calusal delivery}

%Given a succinct reference implementation $SRImp(Spec) = (Q_s,\Sigma_s,q_0^s,\rightarrow_s)$, and implementation $imp = (f_M,f_{apl},d_{\mathit{init}},f_{o},f_{arb})$ and its semantics $Sem(imp) = (Q_{\mathit{imp}},$ $\Sigma_{\mathit{imp}},\rightarrow_{\mathit{imp}},q_{\mathit{imp}}^0)$, a relation $R_{\mathit{sim}}$ is defined as follows: for $q_{\mathit{imp}} = (data,msgs) \in Q_{\mathit{imp}}$ and $q_s = (O_s,\mathit{ro}_s,\mathit{del}_s,\mathit{arb}_s) \in Q_s$, $(q_{\mathit{imp}},q_s) \in R_{\mathit{sim}}$, if: Let $O_{\mathit{imp}} = f_{o}(data,msgs)$, 

Let us take the OR-set algorithms of \cite{Shapiro:2011} as example to define the simulation relation $R_{\mathit{sim}}$. Given a state $q_{\mathit{imp}} = (data,msgs,\mathit{hb})$ of implementation and a state $q_s = (O,\mathit{ro},\mathit{del},\mathit{arb})$ of specification, $(q_{\mathit{imp}},q_s) \in R_{\mathit{sim}}$ often has the following requirements: 

$\newline$ \noindent - A inductive invariants for both $q_{\mathit{imp}}$ and $q_s$. The inductive invariants normally states some features of local states and operations for executions starting from initial state.

Example: The inductive invariant of $q_{\mathit{imp}}$ requires that  

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] For each $id = (r,c)$ in $S \cup msgs$, $c$ is less or equal than counter of replica $r$. 

    \item[-] If $(a,\mathit{Set},r)$ is in message, then for each $id \in \mathit{Set}$, $(a,id) \notin data[r].S \wedge (a,id,\_,r) \notin msgs$. 
    \end{itemize} 
    
    The inductive invariants of $q_s$ requires that, each $rem$ operation is in $\mathit{FstRem}$ of some $add$ operation. 

$\newline$ \noindent - There exists a map $f$ between $\Sigma_e$ and $O_s$, where $O_s$ is union of $O$ and operations in $\mathit{arb}$. 

Example: $f$ maps $(a,id)$ into $add(a)$ and maps $(a,\mathit{Set},r)$ into $rem(a)$. 

$\newline$ \noindent - We require that $f(f_{arb}(data,msgs))$ be a sub-sequence of $\mathit{arb}$. 

$\newline$ \noindent - For each $x \in \Sigma_e$, there is a message of $x$ waiting to be applied to replica $r$, if and only if $f(x) \notin visTo(O,r,\mathit{vis})$. 

$\newline$ \noindent - There exists a method to check whether an ``$add$ operation'' of $\Sigma_e$ is redundant. If $x$ is redundant, then $f(x) \notin O_s$, and there is no message of $x$ in $msgs$.

Example: $(a,id)$ is redundant, if for each replica $r$, $(a,id) \notin data[r].S$.  

$\newline$ \noindent - Stating the $\mathit{FstRem}$ relations for elements of $\Sigma_e$. 

Example: Given $x=(a,\mathit{Set},r)$: for each $id \in \mathit{Set}$, $f(x) \in \mathit{FstRem}(O,\mathit{vis},f(a,id))$; while for each $id \notin mathit{Set}$, $f(x) \notin \mathit{FstRem}(O,\mathit{vis},f(a,id))$.

$\newline$ \noindent - The visibility for elements of $\Sigma_e$. 

Example: $(a,id) \in data[r].S$, if and only if $f((a,id)) \in visTo(O,r,\mathit{vis})$, and for each $o_r \in \mathit{FstRem}(O,\mathit{vis},f((a,id)))$, $o_r \notin visTo(O,r,\mathit{vis})$. 

$\newline$ 

Given implementation state $q_{\mathit{imp}}$, specification state $q_s$ and a implementation transition label $\alpha$, where $(q_{\mathit{imp}},q_s) \in R_{\mathit{sim}}$ and $f$ is the map between $\Sigma_e$ and operations of $q_s$. The function $f_s$ of $f_s$-simulation is defined as follows: 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alpha = apply(m)$, $m$ is a message of $x \in \Sigma_e$ and $x$ is redundant, then $f_s(q_{\mathit{imp}},q_s,\alpha) = \epsilon$. 

\item[-] If $\alpha = apply(m)$, $m$ is a message of $x \in \Sigma_e$ and $x$ is not redundant, then $f_s(q_{\mathit{imp}},q_s,\alpha) = addDel(f(x),r)$, where $r$ is the destination replica of $m$. 

\item[-] If $\alpha = (m,a,b,r,\mathit{arb}_{\mathit{imp}})$, then $f_s(q_{\mathit{imp}},q_s,\alpha) = (m,a,b,r,\mathit{arb}'_s)$. If $\alpha$ is a query operation or a $\mathit{rem}$ operation, then $\mathit{arb}'_s$ is same as the arbitration order of $q_s$; Else, if $\alpha$ is a $add$ operation, then $\mathit{arb}'_s$ is obtained from the arbitration order of $q_s$ by inserting the $add$ operation at some place. 
\end{itemize} 

To define function $f_t$, we further require that, if $(q_{\mathit{imp}},q_s) \in R_{\mathit{sim}}$, $q_{\mathit{imp}} {\xrightarrow{\alpha}}_{\mathit{imp}} q'_{\mathit{imp}}$, let $f$ is the map for $q_{\mathit{imp}}$ and $q_s$, then,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $f_s(q_{\mathit{imp}},q_s,\alpha) \neq \epsilon$, then $q_s {\xrightarrow{f_s(q_{\mathit{imp}},q_s,\alpha)}}_s q'_s$, $(q'_{\mathit{imp}},q'_s) \in R$. Let $f'$ be the map for $q'_{\mathit{imp}}$ and $q'_s$, then $f'$ and $f$ map same element into same operation. 

\item[-] If $f_s(q_{\mathit{imp}},q_s,\alpha) = \epsilon$, then $f$ is the map for $(q'_{\mathit{imp}},q_s) \in R$. 
\end{itemize}

The function $f_t$ is defined as follows: Given a trace $t = \alpha_1 \cdot \ldots \cdot \alpha_k$ of $Sem(imp)$, $f_t(t)$ works as follows: 

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Starts from the initial state of $SRImp(Spec)$. Assume we have already dealt with $\alpha_1 \cdot \ldots \cdot \alpha_{u-1}$ and reach state $q_{u-1}$ of $SRImp(Spec)$.    

\item[-] When $\alpha_u$

\end{itemize}

$A_2$ refines $A_1$ w.r.t $f_t$, if for each trace $t_1 = \alpha_1 \cdot \ldots \cdot \alpha_k$ of $A_1$, $t_2 = f_t(\alpha_1) \cdot \ldots \cdot f_t(\alpha_1 \cdot \ldots \cdot \alpha_k)$ is an execution of $A_2$.













\forget
{
$(q_{\mathit{imp}},q_s) \in R_{\mathit{sim}}$, if

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $q_{\mathit{imp}}$ is correct, if

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] For each replica $r$, $E(r) \cap T(r) = \emptyset$.

    \item[-] For each $(a,id)$ in message, it has already in $E \cup T$.

    \item[-] The counter is correct.

    \item[-] If $(a,S,r)$ is in message, then for each $id \in S$, $(a,id) \notin E(r) \wedge (a,id,\_,r) \notin msgs$.
    \end{itemize}

\item[-] $q_s$ is always correct.

\item[-] $f$ maps $(a,id)$ into some $add(a)$ operation, and maps $(a,S,r)$ into some $rem(a)$ operation.

\item[-] $P_{\mathit{redt}}(q_{\mathit{imp}},(a,id))$ holds, if for each replica $r$, $(a,id) \in T(r)$.

\item[-] Given $x=(a,S,r)$: for $id \in S$, $f(x) \in Minus(q_s,f(a,id))$; while for $id \notin S$, $f(x) \notin Minus(q_s,f(a,id))$.

\item[-] $(a,id) \in E(r)$, if and only if $f((a,id))$ is seen by $r$ and no of its corresponding remove is seen by replica $r$.

\item[-] $(a,id) 
}























\forget
{
\section{Simulation Relation}
\label{sec:simulation relation}

In this section, we give our simulation relation between a implementation $\llbracket imp \rrbracket$ and a reference implementation (with causal delivery) $A$. For plus-minus specification and its compacted reference implementation $A'$, we give our simulation relation between $\llbracket imp \rrbracket$ and $A'$. For both simulation relations, we prove that they are sound and complete for refinement checking.

Given $\llbracket imp \rrbracket = (Q_{imp},\Sigma_{imp},\rightarrow_{imp},q_{0imp})$ and $RImp(Spec) = (Q_s,\Sigma_s,vis,q_{0s},li,\rightarrow_s,livReq)$, we say that $R \subseteq Q_{imp} \times Q_s$ is a simulation relation, if given $(q_i,q_s) \in R$

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $q_i {\xrightarrow{m(a,b,r)}}_{imp} q'_i$, then $q_s {\xrightarrow{m(a,b,r)}}_s q'_s$ and $(q'_i,q'_s) \in R$.

\item[-] If $q_i {\xrightarrow{apply(m)}}_{imp} q'_i$, then $q_s {\xrightarrow{addDel(o,r)}}_s q'_s$ and $(q'_i,q'_s) \in R$. Here $o$ and $r$ are obtained as follows: Assume $m=(\_,r',r)$ is the $i-th$ among messages of $q_i$ with source replica $r'$ and destination replica $r$ w.r.t $<_{sd}$. Then, $o$ is the $i-th$ among operations of $q_s$ which happens on replica $r'$ and not visible to replica $r$ w.r.t $ro$.
\end{itemize}




\forget{
Given $\llbracket imp \rrbracket = (Q_{imp},\Sigma_{imp},\rightarrow_{imp},q_{0imp})$ (not assuming causal delivery) and $RImp(Spec) = (Q_s,\Sigma_s,vis,q_{0s},li,\rightarrow_s,livReq)$, we say that $R \subseteq Q_{imp} \times Q_s$ is a simulation relation, if given $(q_i,q_s) \in R$ where $q_i = (repD,msgs,<_{roi})$ and $q_s = (uO,ro,del)$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Let $MsgToOp(q_i) = \{ o \vert \exists m = (dat,\_,\_) \in msgs, o=op(dat) \}$ be the set of operations of messages in $q_i$. Let $AOpS(q_s) = \{ o \vert o \in uO, \exists r, \neg (o {\xrightarrow{vis}} r) \}$ be the set of operations in $q_s$ that are not delivered into all replicas. Given $r \in RId$, let $MsgToOp(q_i)(r) = \{ o \vert \exists m = (dat,\_,r) \in msgs, o=op(dat) \}$ and $AOpS(q_s)(r) = \{ o \vert o \in uO, \neg (o {\xrightarrow{vis}} r) \}$. We require that,

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] $\forall r \in RId$, $\vert MsgToOp(q_i)(r) \vert = \vert AOpS(q_s)(r) \vert$.

    \item[-] There is a function $f_{(q_i,q_s)}$ that maps the $i-th$ op of $MsgToOp(q_i)(r)$ w.r.t $<_{roi}$ into the $i-th$ op of $AOpS(q_s)(r)$ w.r.t $ro$ for each $i$. Moreover, $o_1 = OpMsgI(q_i)(r_1)$, $o_2 = OpMsgI(q_i)(r_2)$, $r_1 \neq r_2$, and $o_1$ and $o_2$ are mapped from same $dat$, if and only if $f_{(q_i,q_s)}(o_1)=f_{(q_i,q_s)}(o_2)$. We can see that such $f_{(q_i,q_s)}$ is unique if it exists.
    \end{itemize}

\item[-] If $q_i {\xrightarrow{m(a,b,r)}}_{imp} q'_i$, then $q_s {\xrightarrow{m(a,b,r)}}_s q'_s$ and $(q'_i,q'_s)$.

\item[-] If $q_i {\xrightarrow{apply(m)}}_{imp} q'_i$, $m=(dat,\_,r)$ and $f_{(q_i,q_s)}(op(dat))=o$, then $q_s {\xrightarrow{addDel(o,r)}}_s q'_s$ and $(q'_i,q'_s)$.
\end{itemize}

}




We say that a transition system is deterministic, if for each state $q$ and transition label $\alpha$, from $q$ there is at most one transition with label $\alpha$. {\color {red}It is safe to assume that when constructing reference implementation $RImp(Spec)$, in update operation transitions $q_s {\xrightarrow{m(a,b,r)}}_s q'_s$, we permit only one possible identifier of the newly added operation of $q'_i$. For example, each replica has a counter for generating unique identifier of operations. In this way, it is obvious that $RImp(Spec)$ is deterministic.}

Given a sequence $t_{imp} = \alpha_1 \cdot \ldots \in \llbracket imp \rrbracket$ and a sequence $t_s = \beta_1 \cdot \ldots \in RImp(Spec)$, we say that $t_{imp}$ and $t_s$ correspond, if

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alpha_i$ is an operation, then $\beta_i = \alpha_i$.

\item[-] Else, $\alpha_i = apply(m)$ and $\beta_i = addDel(o,r)$. Here $o$ and $r$ are obtained as follows: Assume after doing $\alpha_1 \cdot \ldots \cdot \alpha_{i-1}$, $m=(\_,r',r)$ is the $i-th$ among ``messages with source replica $r'$ and destination replica $r$ and still not applied'' w.r.t the occurring order of $t_{imp}$. {\color {red}Here we assume that in $\llbracket imp \rrbracket$, the set of newly generated messages could be seen as ghost variable of $m(a,b,r)$ transitions.}

    Then, after doing $\beta_1 \cdot \ldots \cdot \beta_{i-1}$, $o$ is the $i-th$ among ``operations which happens on replica $r'$ and not visible to replica $r$'' w.r.t the occurring order of $t_s$. {\color {red} Here we assume that in $RImp(Spec)$, the newly generated operation could be seen as ghost variable of $m(a,b,r)$ transitions.}
\end{itemize}

We can prove that, given $t_{imp}$, there exists at most one $t_s$, such that $t_{imp}$ and $t_s$ correspond. The reason is that, (1) $RImp(Spec)$ is deterministic, and (2) to deal with $apply(m)$ transition, there is only one possible selection of operation $o$.

We say that $RImp(Spec)$ trace refines $\llbracket imp \rrbracket$, if $\forall t_{imp} \in \llbracket imp \rrbracket$, $\exists t_s \in RImp(Spec)$, such that $t_{imp}$ and $t_s$ correspond. It is obvious that our simulation relation implies trace inclusion. Since $RImp(Spec)$ is deterministic, similarly as that of \cite{Abadi:1991,Lynch:1995}, we could prove that the opposite direction also holds. Therefore, our simulation relation is a sound and complete method to prove trace refinement, as states by the following theorem:

\begin{theorem}
\label{theorem:equivalence of our simulation relation and sequence inclusion}
$RImp(Spec)$ trace refines $\llbracket imp \rrbracket$, if and only if there exists a simulation relation between $\llbracket imp \rrbracket$ and $RImp(Spec)$.
\end{theorem}



When we consider implementation and reference implementation with causal delivery, above definitions and lemmas still hold. The reason is that, we could easily obtain $<_{sd}$ order from $<$ order in state of implementation.




\subsection{Simulation Relation for Plus-Minus Specification}
\label{subsec:simulation relation for plus-minus specification}

Let $\llbracket imp \rrbracket_{pm}$ be obtained from $\llbracket imp \rrbracket$ by making applying of some messages invisible. Formally,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Each state is of the form $(repD,msgs,<_{sd})$. We change $msgs$ such that there are two kinds of messages: $(dat,r,r')$ and $(data,r,r')_{inv}$.

\item[-] $\Sigma = \{ \epsilon, m(a,b,r), apply(msg) \vert m \in M, a,b \in D, r \in RId, msg \in Msg \}$ is the set of transition labels.

\item[-] When doing operation transitions, some message is transformed into invisible. For example, if $(redD,msgs,<_{sd}) {\xrightarrow{m(a,b,r)}} (redD',msgs',<'_{sd})$ is a transition of $\llbracket imp \rrbracket$, then $(redD,msgs,<_{sd}) {\xrightarrow{m(a,b,r)}} (redD',f(msgs'),<'_{sd})$ is a transition of $\llbracket imp \rrbracket_{pm}$. Here $f(msgs')$ is obtained from $msgs'$ by adding subfix $inv$ into some messages.
\item[-] When applying a message $(\_,\_,\_)$, same as before.

When applying a message $(\_,\_,\_)_{inv}$, use $\epsilon$ as transition label.
\end{itemize}


Given $\llbracket imp \rrbracket_{pm} = (Q_{imp},\Sigma_{imp},\rightarrow_{imp},q_{0imp})$ and $CRImp(Spec) = (Q_s,\Sigma_s,vis,q_{0s},li,\rightarrow_s,livReq)$, we say that $R \subseteq Q_{imp} \times Q_s$ is a weak simulation relation, if given $(q_i,q_s) \in R$

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $q_i {\xrightarrow{m(a,b,r)}}_{imp} q'_i$, then $q_s {\xrightarrow{m(a,b,r)}}_s q'_s$ and $(q'_i,q'_s) \in R$.

\item[-] If $q_i {\xrightarrow{apply(m)}}_{imp} q'_i$, $m=(\_,r',r)$, then $q_s {\xrightarrow{addDel(o,r)}}_s q'_s$ and $(q'_i,q'_s) \in R$. Here $o$ is obtained as follows: Assume $m$ is the $i-th$ among messages of $q_i$ with source replica $r'$, destination replica $r$ and are visible w.r.t $<_{sd}$. Then, $o$ is the $i-th$ among operations of $q_s$ which happens on replica $r'$ and not visible to replica $r$ w.r.t $ro$.

\item[-] If $q_i {\xrightarrow{apply(m)}}_{imp} q'_i$, $m=(\_,r',r)_{inv}$, then $(q'_i,q_s) \in R$.
\end{itemize}


Given a sequence $t_{imp} = \alpha_1 \cdot \ldots \in \llbracket imp \rrbracket_{pl}$ and a sequence $t_s = \beta_1 \cdot \ldots \in CRImp(Spec)$, we say that $t_{imp}$ and $t_s$ correspond, if

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alpha_i$ is an operation, then $\beta_i = \alpha_i$.

\item[-] Else, if $\alpha_i = apply(m=(\_,r',r))$ and $\beta_i = addDel(o,r)$. Here $o$ is obtained as follows: Assume after doing $\alpha_1 \cdot \ldots \cdot \alpha_{i-1}$, $m$ is the $i-th$ among ``messages with source replica $r'$ and destination replica $r$ and still not applied and are visible'' w.r.t the occurring order of $t_{imp}$. Then, after doing $\beta_1 \cdot \ldots \cdot \beta_{i-1}$, $o$ is the $i-th$ among ``operations which happens on replica $r'$ and not visible to replica $r$'' w.r.t the occurring order of $t_s$.

\item[-] Else, $\alpha_i = apply(m=(\_,r',r)_{inv})$ and $\beta_i = \epsilon$.
\end{itemize}

We say that $CRImp(Spec)$ weak trace refines $\llbracket imp \rrbracket_{pl}$, if $\forall t_{imp} \in \llbracket imp \rrbracket_{pl}$, $\exists t_s \in CRImp(Spec)$, such that $t_{imp}$ and $t_s$ correspond. It is obvious that our weak simulation relation implies trace inclusion. Since $RImp(Spec)$ is deterministic, similarly as that of \cite{Abadi:1991,Lynch:1995}, we could prove that the opposite direction also holds. Therefore, our weak simulation relation is a sound and complete method to prove trace refinement, as states by the following theorem:

\begin{theorem}
\label{theorem:equivalence of our simulation relation and sequence inclusion for add-plus specification and its compacted reference implementation}
$CRImp(Spec)$ trace refines $\llbracket imp \rrbracket_{pl}$, if and only if there exists a weak simulation relation between $\llbracket imp \rrbracket_{pl}$ and $CRImp(Spec)$.
\end{theorem}
}

